---
title: "Question 1"
author: "Shaheer Ahmad"
date: "2024-10-02"
output: html_document


#Q1. When is the best time of day, day of the week, and time of year to fly to minimise delays?

#Importing the required libraries
```{r}
library(data.table)#Importing the libraries
library(data.table)
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
```

#Opening the datasets 
```{r}
d_26<- fread("C:/Users/haysh/Downloads/2006.csv")
d_27<- fread("C:/Users/haysh/Downloads/2007.csv")
```{r}
```
#Finding Column Names
```{r}
colnames(d_26)
```

#Getting an understanding of the variables of the 
```{r}
head(d_27,10)
```

#Filtering the required columns to answer the question
```{r}
d_26 %>% select("Year","Month","DayofMonth","DayOfWeek","DepTime","DepDelay")
d_27 %>% select("Year","Month","DayofMonth","DayOfWeek","DepTime","DepDelay")`

```
#Combining the two datasets,transforming into data frame and getting the first 10 rows

```{r}
flight_ <- rbind.data.frame(d_26,d_27)
flight_delay <-data.frame(flight_)
head(flight_delay,10)
```
#Transforming the ArrivalTime column into hour format as a new column called 'ArrivalTime'
```{r}
flight_delay$ArrTime <- flight_delay$ArrTime/100
head(flight_delay,10)
df <- data.frame(flight_delay)
```
#Dropping the null values

```{r}
df <- na.omit(df)
```
#As there are Arrival Times greater than 24.00 which is of no value, filtering only the rows which the Arrival Time is less than 24.00
```{r}
`df <- df %>% filter(ArrTime <= 24.00)
options(max.print=999999999)``{r}
```
#Best time of the day?
#Filtering Arrival Time and ArrDelay columns
```{r}
df1 <- df[ , c("ArrTime","ArrDelay")] 
head(df1,10)
```
#Creating a new column which has transformed the hours of day into four time slot bins.

```{r}
breaks <- c(0.00,6.00,12.00,18.00,23.59)
tags <- c("Early Morning","Morning","Afternoon","Evening")
```
# bucketing values into bins

```{r}
group_tags <- cut(df$ArrTime, 
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags)
```
# inspect bins

```{r}
summary(group_tags)
```
#Replacing the new column with 'ArrTime' column

```{r}
df['ArrTime']<-group_tags
summary(df)
```

#Grouping the Arrival delays by Timezone bins and obtaining the mean
```{r}
delays_byTimeZone <- df %>% group_by (ArrTime) %>% summarise(Arr_delay = mean(ArrDelay),.groups = 'drop')
delays_byTimeZone =data.frame(delays_byTimeZone)
delays_byTimeZone %>% drop_na()
is.null(delays_byTimeZone) 
```
dropping the null values and converting getting the summary
```{r}
delays_byTimeZone %>% drop_na()
is.null(delays_byTimeZone) 
summary()
```
#Plotting the findings

```{r}
barplot(delays_byhour1$Arr_delay,names.arg =delays_byhour1$ArrTime )
```

#Finding the median values and getting the summary
```{r}
delays_byTimeZone1 <- df %>% group_by (ArrTime) %>% summarise(Arr_delay = median(ArrDelay),.groups = 'drop')
delays_byTimeZone1 =data.frame(delays_byTimeZone1)
delays_byTimeZone1 %>% drop_na()
is.null(delays_byTimeZone1)
summary() 
```
#BoxPlot
```{r}
boxplot(df$ArrDelay)
```{r}
```
#Plotting the findings

```{r}
barplot(delays_byTimeZone1$Arr_delay,names.arg =delays_byTimeZone1$ArrTime )
```

#Best time of the week?
#Filtering Arrival Time and ArrDelay columns

```{r}
df2 <- df[ , c("DayOfWeek","ArrDelay")] 
options(max.print=9999999)
df2
```
#Grouping the Arrival delays by Day of Week and obtaining the mean

```{r}
delays_byday <- df %>% group_by (DayOfWeek) %>% summarise(Arr_delay = mean(ArrDelay),.groups = 'drop')
delays_byday =data.frame(delays_byday)
is.null(delays_byday$ArrTime)
is.null(delays_byday)
delays_byday %>% drop_na()
summary()
```
#Plotting the findings
```{r}
barplot(delays_byday$Arr_delay,names.arg =delays_byday$DayofWeek,xlab='Day of Week',ylab=
          'Mean delay')
```

##Filtering Month and ArrDelay columns
```{r}
delays_byday %>%
  ggplot(aes(x=delays_byday$DayofWeek,y=delays_byday$ArrDelay)) +
  geom_point(size = 3, colour = "black") + 
  geom_segment( aes(x= delays_byday$DayofWeek, xend= delays_byday$DayofWeek, y= 0, yend=delays_byday$ArrDelay))+
  labs(y= "Mean Arrival Delay", x="=Day")

```

#Grouping the Arrival delays by Day of Week and obtaining the median

```{r}
delays_byday1 <- df %>% group_by (DayOfWeek) %>% summarise(Arr_delay = median(ArrDelay),.groups = 'drop')
```
#Plotting the findings

```{r}
barplot(delays_byday1$Arr_delay,names.arg =delays_byday1$DayofWeek,xlab='Day of Week',ylab='Median delay')
```
#Best time of the year?

```{r}
df3 <- df[ , c("Month","ArrDelay")] 
options(max.print=999999999)
```
#Grouping the Arrival delays by Month and obtaining the mean

```{r}
delays_by_day_of_year <- df %>% group_by (Month) %>% summarise(Arr_delay = mean(ArrDelay),.groups = 'drop')
delays_by_day_of_year=data.frame(delays_by_day_of_year)
delays_by_day_of_year %>% drop_na()
```
#Plotting the findings

```{r}
plot(delays_by_day_of_year$Arr_delay,names.arg =delays_by_day_of_year$Month,xlab='Month',ylab=
          'Mean delay')
```
#Grouping the Arrival delays by Month and obtaining the median

```{r}
delays_by_day_of_year1 <- df %>% group_by (Month) %>% summarise(Arr_delay = median(ArrDelay),.groups = 'drop')
delays_by_day_of_year1=data.frame(delays_by_day_of_year1)
delays_by_day_of_year1 %>% drop_na()
```
#Plotting the findings

```{r}
ggplot(delays_by_day_of_year, aes(x=delays_by_day_of_year$Month, y=delays_by_day_of_year$Arr_delay)) +
  geom_line()

ggplot(delays_by_day_of_year, aes(x=delays_by_day_of_year1$Month, y=delays_by_day_of_year1$Arr_delay)) +
  geom_line()
```
  
