---
title: "QUESTION 2"
author: "Shaheer Ahmad"
date: "2024-03-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Q2. Do older planes suffer more delays?

#Importing the required libraries

```{r}
library(data.table)
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
```

#Opening the datasets and combining them

```{r}
d_27<- fread("C:/Users/haysh/Downloads/2007.csv")
d_26<- fread("C:/Users/haysh/Downloads/2006.csv")
flight_ <- rbind.data.frame(d_26,d_27)
```
#converting as a dataframe

```{r}
delay_ <- data.frame(flight_)
head(delay_,10)
```
#Filtering reqiuired columns of the dataset

```{r}
delay_ %>% select("ArrDelay","TailNum","Year")
head(delay_,10)
```

#Opening the Planes Dataset

```{r}
planes <-fread("C:/Users/haysh/Downloads/plane-data_1.csv",header = TRUE,fill=TRUE)
```
#Converting as a dataframe

```{r}
planes<- data.frame(planes)
head(planes,10)
```
#Filtering the required Columns 

```{r}
planes <- planes[ , c("tailnum","year")]
options(max.print=9999999)
head(planes,10)
```
#Filtering inorder to obtain records of delays

```{r}
delay_ <- delay_ %>% filter(ArrDelay > 0)
```

#Dropping Null Values

```{r}
delay_ %>% drop_na()
planes %>% drop_na()
```
#Making a column column name inorder to merge the two datasets
   
```{r}
colnames(delay_)[11] <- "tailnum"
head(delay_,10)
head(planes,10)
colnames(planes) <- c('tailnum')
colnames(planes)[1]  <- "tailnum" 
```
#Merging the two data sets using the common column

```{r}
options(max.print=9999999)
merged <- merge(delay_,planes, by ="tailnum")
head( merged,10)
```
#Dropping null values and doing necessary data wrangling 
```{r}
merged %>% drop_na()
head(merged,10)
colnames(merged)[30]  <- "ManufacturedYear" 
head(merged,10)
options(max.print=9999999)
merged %>% drop_na()
sumary(merged,10)
```

#Converting year and manufactured year variables as integers, and calculating the age of each plane
```{r}
merged$Year<-as.integer(merged$Year)
merged$ManufacturedYear<-as.integer(merged$ManufacturedYear)
Age <- merged$Year - merged$ManufacturedYear
merged$Age <- Age
head(merged,10)
options(max.print=9999999)
merged %>% drop_na()
head(merged,10)
```

#Categorizing age column values into bins 
```{r}
breaks <- c(0,10,20,30,40,51)
tags <- c('10 years<=','20 years<=','30 years<=','40 years<=','51 years<=')
# bucketing values into bins
group_tags <- cut(merged$Age, 
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags)
```

```{r}
# inspect bins
summary(group_tags)
merged['Age']<-group_tags
summary(merged)
merged %>% drop_na()
head(merged,10)
```
                  
#Calculating the group mean arrival delay values of each group
```{r}
merged_by_age <- merged %>% group_by (Age) %>% summarise(Arr_delay = mean(ArrDelay),.groups = 'drop')
merged_by_age =data.frame(merged_by_age)
merged_by_age %>% drop_na()
summary(merged_by_age)

```
     
#Plotting the findings             
```{r}
#Plotting the findings
barplot(merged_by_age$Arr_delay,names.arg =merged_by_age$Age ) 


```
#Calculating the group median arrival delay values of each group
```{r}
merged_by_age1 <- merged %>% group_by (Age) %>% summarise(Arr_delay = median(ArrDelay),.groups = 'drop')
merged_by_age1 =data.frame(merged_by_age)
merged_by_age1 %>% drop_na()
summary(merged_by_age1)

```
#Plotting the findings

```{r}
barplot(merged_by_age1$Arr_delay,names.arg =merged_by_age$Age ) 
a<-merged$ArrDelay
b<-merged$ManufacturedYear
cor(a,b)
```
                  
