---
title: "QUESTION 3"
author: "Shaheer Ahmad"

output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Importing the libraries
```{r}
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)

```

#Opening the Data Sets

```{r}
D_27<- fread("C:/Users/haysh/Downloads/2007.csv")
D_26<- fread("C:/Users/haysh/Downloads/2006.csv")
head(D_26,10)
```
#Filtering the necessary columns

```{r}
d_26<-D_26 %>% select('Year','Month','FlightNum','Origin')
d_27<-D_27 %>% select('Year','Month','FlightNum','Origin')
head(d_26,10)
```
#Removing Null Values

```{r}
d_26 %>% drop_na()
d_27 %>% drop_na()
head(d_26,10)
head(d_27,10)
```
#Categorizing the 12 months of the year into quarters

```{r}
breaks <- c(1,3,6,9,12)
tags <- c('Quarter1','Quarter2','Quarter3','Quarter4')
# bucketing values into bins
group_tags1 <- cut(d_26$Month,
                  breaks=breaks, 
                  include.lowest=FALSE, 
                  right=FALSE, 
                  labels=tags)
summary(group_tags1)

#Replacing the values of 'Month' column with categorized values
d_26$Month<-group_tags1
head(d_26,10)
breaks <- c(1,3,6,9,12)
# specify interval/bin labels
tags <- c('Quarter1','Quarter2','Quarter3','Quarter4')
# bucketing values into bins
group_tags2 <- cut(d_27$Month,
                  breaks=breaks, 
                  include.lowest=FALSE, 
                  right=FALSE, 
                  labels=tags)
d_27$Month<-group_tags2
d_27 %>% drop_na()
d_26 %>% drop_na()
```
#Grouping and taking the counts using Flight number variable
```{r}
a<-count(d_26,Month,Year)
a =data.frame(a)
head(a,10)
a <-a[-c(9,10),]
a1<-a[-5,]
a1 %>% drop_na()
a1
b<-count(d_27,Month,Year)
b =data.frame(b)
b <-b[-c(9,10),]
b1<-b[-5,]
b1 %>% drop_na()
b1
```

```{r}
flight_ <- rbind.data.frame(a1,b1)
head(flight_,10)
options(max.print=9999999)
barplot(flight_$n, main="No of Flights in each Quarter",
        xlab="=X axis - Quarter1-2006,Quarter1-2007,Quarter2-2006,Quarter2-2006,Quarter2-2007,Quarter3-2006,Quarter3-2007,Quarter4-2006,Quarter4-2007")

```
#Location wise Analysis
#create a dataframe filtered for the top 20 Airports, grouped by Month and Year

  

```{r}
location <- combined %>%
  group_by(Origin) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  head(56) %>%
  pull(Origin)
```


```{r}
 location10 <- combined %>%
  filter(Origin %in% location)
```

#Finding the number of flights conceded in each airport in each quarter of year 2006 and 2007
```{r}
location10 <- location10 %>%
  group_by(Origin, Year, Month) %>%
  summarise(Count = n()) %>%
  ungroup()
```
#Sorting values according to count order for each year
```{r}
location10 %>%
  arrange(Count) %>%
  head(80)
```

```{r}
location11 <- combined %>%
  filter(Origin %in% location)
```


```{r}
location11 <- location11 %>%
  group_by(Origin) %>%
  summarise(Count = n()) %>%
  ungroup()
```

#Finding the values of 80 airports which have the most number of flights concededs
```{r}
location11 %>%
  arrange(Count) %>%
  head(80)
```

```{r}
top_airports <- combined %>%
  count(Origin) %>%
  arrange(desc(n)) %>%
  head(20) %>%
  pull(Origin)
```

#filter only for the top 10 airports
```{r}
top_10 <- combined %>%
  filter(Origin %in% top_states)
```

```{r}
top_10 <- top_10 %>%
  group_by(Origin, Year, Month) %>%
  summarise(Count = n()) %>%
  ungroup()

```
#Plotting the final result
```{r}

side <- ggplot(top_10, aes(x = Origin, y = Count, fill = as.factor(Year))) +
  geom_bar(stat = "identity", position = "dodge") +
 
  labs(x = "State", y = "Number of Flights", title = "Number of Flights over time by States") +
  theme(plot.title = element_text(size = 18),
        axis.text.x = element_text(size = 13),
        axis.text.y = element_text(size = 13),
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 1, title.position = "top"))

```


