---
title: "QUESTION 4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Q4

#Importing the required libraries
```{r}
library(data.table)#Importing the libraries
library(data.table)
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
```
#Opening the datasets and merging them

```{r}
```


```{r}
d_26<- fread("C:/Users/haysh/Downloads/2006.csv")
d_27<- fread("C:/Users/haysh/Downloads/2007.csv")
flight_ <- rbind.data.frame(d_26,d_27)
combined <-data.frame(flight_)
head(combined,10)
```


```{r}
```
#Converting to datetime
#extracting the minutes and hours

```{r}
Min1 <-  function(x){
  substr(x, nchar(x)-1, nchar(x))
}
Min1 <- as.integer(Min1(combined$CRSDepTime))

Hr <- floor(combined$CRSDepTime / 100)
head(combined,10)
```
#converting to datetime

```{r}
combined <- mutate(combined,CRSDepTime = make_datetime(Year, Month,DayofMonth,Hr, Min1)) 
head(combined,10)    
```
#Sorting the dataset according to CRSDeptime

```{r}
combined[order(combined$TailNum,combined$CRSDepTime, decreasing = TRUE), ]
head(combined,10)

```
#Defining the new variable which is the sum of both Arrival and Departure delays
```{r}
combined$TotalCDelay <- combined$DepDelay + combined$ArrDelay
head(combined,10)
```
#Lagging the current delay with one time period inorder to find previous delays
```{r}
combined <- combined %>%
  mutate(Delay1 = lead(ArrDelay))
head(combined,10)
head(combined,10)

```
#Dropping the null values
```{r}
combined <- na.omit(combined)
```
#plotting the cascading delays
```{r}
ggplot(data = combined, aes(x = combined$TotalCDelay, y = combined$Delay1)) +
  geom_point(color = "#83D6D1") +
  geom_smooth(se = FALSE, color = "#2B229A") +
  ggtitle("Cascading delays") +
  labs(x= "TotalCurrentDelay", y = "Delay1") +
  theme_classic()
```
#CrossTabulation
```{r}
combined$TotalCDelay1 <- ifelse(combined$TotalCDelay>0, 1,0)
combined$Delay11 <- ifelse(combined$Delay1>0, 1, 0)

crosstab <- table(combined$TotalCDelay1, combined$Delay11)


```
#Plotting a scatterplot using LateAircraftDelay and ArrDelay variables as x axis and y axis

```{r}
ggplot(data = combined, aes(x = LateAircraftDelay, y =ArrDelay)) +
  geom_point(color = "#83D6D1") +
  geom_smooth(se = FALSE, color = "#2B229A") +
  ggtitle("Cascading delays") +
  labs(x= "TotalCDelay", y = "Delay1") +
  theme_classic()
```
  

```{r}
```
