---
title: "Question 5"
author: "Shaheer"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Importing the libraries
```{r}
library(data.table)
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
library(caTools)
install.packages('corrgram')
library(corrgram)
```
Opening the datasets
```{r}
d_26<- fread("C:/Users/haysh/Downloads/2006.csv")
d_27<- fread("C:/Users/haysh/Downloads/2007.csv")
colnames(d_26)
```

#Combining the two datasets
```{r}
flight_ <- rbind.data.frame(d_26,d_27)
flight_delay <-data.frame(flight_)
head(flight_delay,10)
```
#Transforming the ArrivalTime column into hour format as a new column called 'ArrivalTime'

```{r}
flight_delay$ArrTime <- flight_delay$ArrTime/100
head(flight_delay,10)
df <- data.frame(flight_delay)
```

#Dropping the null values

```{r}
df <- na.omit(df)
head(df,10)
```
#As there are Arrival Times greater than 24.00 which is of no value, filtering only the rows which the Arrival Time is less than 24.00

```{r}
df <- df %>% filter(ArrTime <= 24.00)
options(max.print=999999999)
head(df,10)
```
#Creating a new column which has transformed the hours of day into four time slot bins.

```{r}
breaks <- c(0.00,6.00,12.00,18.00,23.59)
tags <- c("0","1","2","3")
# bucketing values into bins
group_tags <- cut(df$ArrTime, 
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags)
# inspect bins
summary(group_tags)
```
#Replacing the new column with 'ArrTime' column

```{r}
df['ArrTime']<-group_tags
head(df,10)
summary(df)
```

#Taking a sample of 65000
```{r}
df <- df[sample(nrow(df), 65000),]
summary(df)
class(df)

```
#Making categorical features factors 


```{r}
df$CRSDepTime  <- as.factor(df$CRSDepTime)
df$Month  <- as.factor(df$Month)
df$DayOfWeek  <- as.factor(df$DayOfWeek)
df$UniqueCarrier  <- as.factor(df$UniqueCarrier)
df$Origin  <- as.factor(df$Origin)
df$Dest  <- as.factor(df$Dest)
df$CancellationCode  <- as.factor(df$CancellationCode)
df$Dest  <- as.factor(df$Dest)
head(df,10)
```

#PLotting the correlation matrix
```{r}
corrgram(df,lower.panel = panel.shade,upper.panel=panel.cor)
```
#Selecting the required variables for the model after checking the correlation matrix
```{r}
df<-df %>% select('ArrDelay','CRSDepTime','Month', 'ArrTime','DayofMonth', 'DayOfWeek', 'CarrierDelay','WeatherDelay','NASDelay','LateAircraftDelay')
head(df,10)
```


```{r}
#Encoding the target value
df <- df %>%
mutate(delayed = if_else(ArrDelay > 0, 1 , 0))
```

#Plotting boxplots in order to understand the distributions of the variables
```{r}
boxplot(y=df$DepDelay,df$DepDelay,main ="DepDelay")

boxplot(y=df$DepDelay,df$DepDelay,main ="DepDelay")
boxplot(y=df$CRSArrTime,df$CRSArrTime,main ="CRSArrTime")
boxplot(y=df$CRSElapsedTime,df$CRSElapsedTime,main ="CRSElapsedTime")
boxplot(y=df$CRSElapsedTime,df$CRSElapsedTime,main ="CRSElapsedTime")

```
#Splitting the dataset as trainset and test set
```{r}
`df$delayed <- as.factor(df$delayed)
testsplit<- sample.split(Y=df$ArrDelay,SplitRatio<-0.7)
trainData <- subset(x=df,testsplit==TRUE)
head(trainData)
testData <- subset(x=df,testsplit==FALSE)
```

#Fitting the model
```{r}
model<-lm(ArrDelay~.,data=trainData)

summary(model)
```

#Plotting the residuals in order to check whether its been normally distributed or not
```{r}
res<- as.data.frame(residuals(model))
ggplot(res, aes(residuals(model)))+
  geom_histogram(fill='blue')
pred<-predict(model,testData)
modelvals<- cbind(testData$Weight,pred)
colnames(modelvals)
```
#Calculating the R squared value
```{r}
library(caret)

r2 <- R2(y_pred, y_test)
```



#Randomforrest regressor
```{r}
library(randomForest)

Rf <- randomForest(X_train, y_train, ntree = 15, mtry = ncol(X_train),
                   replace = FALSE, sampsize = nrow(X_train), nodesize = 1,
                   importance = TRUE)

Rf_train_pred <- predict(Rf, X_train)
Rf_test_pred <- predict(Rf, X_test)
```

#Evaluating the model by introducing a function
```{r}

def evaluation(model):
    modelA.fit(X_train, y_train)
    y_pred = model.predict(X_test)
    print("R squared: ", round(r2,3))
    print("RMSE: {}".format(np.sqrt(metrics.mean_absolute_error(y_test, y_pred))))
```
    
```{r}
    modelA = Rf
evaluation(Rf)
```

#DecisionTreeRegressor
```{r}
from sklearn.tree import DecisionTreeRegressor
dtm = DecisionTreeRegressor(max_depth=4,
                           min_samples_split=50)
```

```{r}
modelA = dtm
evaluation(dtm)
```
